---
title: 'Part 2: Basic Inferential Data Analysis'
author: "Olha Klishchuk"
date: "2023-10-18"
output: pdf_document
---

# Overview

```{r}
library(datasets)
suppressPackageStartupMessages(library(dplyr))
```

Data\
A data frame with 60 observations on 3 variables.

[,1]	len	    numeric	 Tooth length. \
[,2]	supp	factor	 Supplement type (vitamin C or other J). \
[,3]	dose	numeric	 Dose in milligrams/day. 

```{r}
data(ToothGrowth)
head(ToothGrowth)
```

# Description

```{r}
# Descriptive Statistics and EDA
summary(ToothGrowth)
```
```{r}
df<-ToothGrowth
options(repr.plot.height = 6)
par(mfcol = c(1,2))
with(df, hist(len, breaks = 10, main = "Histogram of tooth length", 
              xlab = "tooth length", ylab = "Frequency", col = "lightblue"))
with(df, hist(dose, xlab = "dose", ylab = "Frequency", col = "lightblue"))
```
 __*Plot 2.*__ Distribution of data by key experimental variables

So from descriptive analysis, which you can find in appendix 2.1, we find out that mean effect from vitamin C injection was 16.96 cm of teeth growth, while in other treatment we obtained - 20.66 cm. Lets test hypothesis that differences between these two treatments are significant.

# Hypothesis Testing

```{r}
# Hypothesis test Ha: m2-m1>0
# Paired test by supplement used
t.test(df[df[['supp']]=='OJ','len'], df[df[['supp']]=='VC','len'], paired = T)$ a
```

```{r}
#Grouping data by dosage
dose_0.5v1<-
df%>%
 group_by(dose)%>%
 mutate(dose=dose, supp = supp, len = len)%>%
 filter(dose %in% c(.5,1))

dose_0.5v2<-
df%>%
 group_by(dose)%>%
 mutate(dose=dose, supp = supp, len = len)%>%
 filter(dose != 1)

dose_1v2<-
df%>%
 group_by(dose)%>%
 mutate(dose=dose, supp = supp, len = len)%>%
 filter(dose > .5)
```

```{r}
# Hypothesis test Ha: B!=0
# Paired test by dosage
data_test<- 
data.frame(statistics=c(t.test(len~dose, dose_0.5v1, paired = F)$statistic,
           t.test(len~dose, dose_0.5v2, paired = F)$statistic,
           t.test(len~dose, dose_1v2, paired = F)$statistic), 
p.value = c(t.test(len~dose, dose_0.5v1, paired = F)$p.value,
           t.test(len~dose, dose_0.5v2, paired = F)$p.value,
           t.test(len~dose, dose_1v2, paired = F)$p.value), 
           row.names = c('0.5v1', '0.5v2','1v2'))
data_test
```
Therefore, the type of supplement is used in treatment appeared to be significant and level of efficiency is tooth growth on 3.7 cm more then on vitamin C using.

```{r}
#paired T-confidence intervals
round(t.test(df[df[['supp']]=='OJ','len'], df[df[['supp']]=='VC','len'], paired 
             = FALSE, conf.level = .95, var.equal = TRUE)$conf, 5)
```

# Summary


- using other vitamins (A, D, E, alpha tocopherol) the tooth growth in guinea pigs were more significant then under solely supplementing vitamin C.
- The mean difference in about 3.7 cm.
- the standard deviation in observation is lower under more  diet on -1.6605


# References
- Crampton, E. W. (1947). The growth of the odontoblast of the incisor teeth as a criterion of vitamin C intake of the guinea pig. The Journal of Nutrition, 33(5), 491–504. doi:10.1093/jn/33.5.491.

# Appendix

## B 

### B.1 Descriptives


```{r}
df<-ToothGrowth
mean_vc<-
    df%>%
    group_by(supp)%>%
    filter(supp == 'VC')%>%
    mutate(mean_len = mean(len), mean_dose = mean(dose))%>%
    select(-c(len, dose))

sd_vc<-
    df%>%
    group_by(supp)%>%
    filter(supp == 'VC')%>%
    mutate(sd_len = sd(len), sd_dose = sd(dose))%>%
    select(-c(len,dose))
cat('mean: ',as.numeric(mean_vc[1,]), 'sd:', as.numeric(sd_vc[1,]))
```
```{r}

mean_oj<-
    df%>%
    group_by(supp)%>%
    filter(supp != 'VC')%>%
    mutate(mean_len = mean(len), mean_dose = mean(dose))%>%
    select(-c(len,dose))

sd_oj<-
    df%>%
    group_by(supp)%>%
    filter(supp != 'VC')%>%
    mutate(sd_len = sd(len), sd_dose = sd(dose))%>%
    select(-c(len,dose))
cat('mean:', as.numeric(mean_oj[1,]), 'sd: ',  as.numeric(sd_oj[1,]))
```

### B.2 Hypothesis test 
$H_a: B!=0$
##### Paired test by dosage     

```{r}
t.test(len~dose, dose_0.5v1, paired = F)

t.test(len~dose, dose_0.5v2, paired = F)

t.test(len~dose, dose_1v2, paired = F)
```

